---
title: "Assignment 4"
Name: Marcus Ramirez
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
```

```{r}
library(tidyverse)
```

# Data Visualization Article Notes:
```{r}
ggplot2::mpg
```
-data frame found in ggplot2 (a rectangular collection of variables & observations)
-mpg contains observations collected on diff. models of cars

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg)
```
-plot mpg & specify what's on each axis
-ggplot(data=mpg) creates an empty graph but u have to add layers
-geom_point() adds layer of points to plot (creates scatterplot)

CODE PRESET:
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>)) 
-replace bracketed sections in code below w/ dataset
-aesthetic: visual property of objects in plot (size, shape, color, etc.) -> referred to as level

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
-maps the colors of each point to the class variable (reveals class of each car)

```{r}
# Left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
-mapping class to alpha controls transperancy of points
-mapping class to shape controls shape of points
-you use aes() to associate the name of the aesthetic with a variable to display
-+ signs have to be at end of line not the beginning (check above)

-facets: sunplots that each display one subset of data
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
-add facet_grid() to plot on combination of 2 variables
-use a . instead of a variable name if u don't wanna facet in row/column dimension

-geom: geometrical object that plot uses to represent data
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
-lets you display multiple geoms in same plot

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```
-allows you to change colors on bar graphs

-coord_flip() switches the x and y axes -> useful for horizontal box plots


# Graphics for Communication Article Notes:
-Label, Annotations, Scales
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```
-labs() function lets you add labels (used to add title here)^
-plot title should summarize main finding instead of saying vs
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```
-labs() also let's you replace axis & legend titles

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
```
-pulls out most efficient car in each class w/ dplyr then labels it
-text makes it a little messy with the overlap
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```
-using label puts rectangles behind text
``{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
```
-not working but ggrepel package automatically adjusts labels to remove overlap

```{r}
label <- mpg %>%
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \nrelated to decreasing fuel economy."
  )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
-computes maximum value of x and y

```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
-moves text exactly on borders of plot using +Inf and -Inf

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))
```
-ggplot2 atuomatically adds scales for you
-breaks controls position of ticks or values associated w/ keys
-labels control text associated w/ each tick/key
-breaks & lablels can also control appearance of legends
-breaks can be used when there's relatively little data & want to highlight exactly where observations occur

```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_x_date(NULL, breaks = presidential$start, date_labels = "'%y")
```

-theme() setting can be used to control overall position of legend
```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") # the default
```
-legend.position = "none" would remove the legend

replacing a scale:
```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d()

ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_bin2d()
```
-here a log transformation performed
-slight problem: axis labels changed
```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
```
-axis now have original labels

```{r}
ggplot(mpg, mapping = aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth() +
  coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))

mpg %>%
  filter(displ >= 5, displ <= 7, hwy >= 10, hwy <= 30) %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()
```
-3 ways to control plot limits
 -adjusting what data plotted
 -setting limits in each scale
 -setting xlim & ylim in coord_cartesian()
 
themes:
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_bw()
```
-allows you to customize non-data elements in plot w/ a theme